---
title: "Summarising results from predictNMB"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summarising-results-with-predictNMB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(predictNMB)
library(parallel)
library(ggplot2)
set.seed(42)
```

# Making plots with predictNMB

This vignette is purely about how to use the `plot()` method and `make_summary_table()` to visualise and summarise the simulations made using `{predictNMB}`. For an introduction to `{predictNMB}`, please see the other vignette.

Firstly, as an example case, we will run `screen_simulation_inputs()`.

```{r}
get_nmb_sampler_training <- function() {
  c(
    "TP" = rnorm(n = 1, mean = -80, sd = 5),
    "TN" = 0,
    "FP" = -20,
    "FN" = rnorm(n = 1, mean = -100, sd = 10)
  )
}

get_nmb_sampler_evaluation <- function() {
  c(
    "TP" = -80,
    "TN" = 0,
    "FP" = -20,
    "FN" = -100
  )
}
```

```{r, eval=FALSE}
cl <- makeCluster(detectCores())
```

```{r, include=FALSE}
chk <- Sys.getenv("_R_CHECK_LIMIT_CORES_", "")
if (nzchar(chk) && chk == TRUE) {
  ncores <- 2
} else {
  ncores <- detectCores()
}
cl <- makeCluster(ncores)
```

## Plotting the results from `screen_simulation_inputs()` 

```{r}
sim_screen_obj <- screen_simulation_inputs(
  n_sims = 500,
  n_valid = 10000,
  sim_auc = seq(0.7, 0.95, 0.05),
  event_rate = c(0.1, 0.2),
  fx_nmb_training = get_nmb_sampler_training,
  fx_nmb_evaluation = get_nmb_sampler_evaluation,
  cutpoint_methods = c("all", "none", "youden", "value_optimising"),
  cl = cl
)
stopCluster(cl)
```

### Choosing the x-axis variable and other constants

In this simulation screen, we vary both the event rate and the model discrimination.
There are many ways that we could visualise these data. The `plot()` allows us to make some basic plots to compare different cutpoint methods regardin the Net Monetary Benefit (NMB) and another variable of our choice. In our case, we can choose between the `sim_auc` and `event_rate`. We control this with the `x_axis_var` argument.

```{r}
plot(sim_screen_obj, x_axis_var = "sim_auc")
plot(sim_screen_obj, x_axis_var = "event_rate")
```

For these plots, one of the screened inputs will be the x-axis variable and the other will only displayed at one of it's levels. The default will assume the first level, so when we visualised the change in NMB as `sim_auc` increased, we were only observing this for the case where `event_rate = 0.1`. We can choose to select another level with the `constants` argument. This argument expects a named list containing the values to keep for the screened inputs which are not shown on the x-axis.

```{r}
plot(sim_screen_obj, x_axis_var = "sim_auc", constants = list(event_rate = 0.1))
plot(sim_screen_obj, x_axis_var = "sim_auc", constants = list(event_rate = 0.2))
```

We see both a change to the plot as well as the message produced when the plot is made.

### Choosing a y-axis variable

There are three options for the y-axis. The default is the NMB, but you can also visualise the Incremental Net Monetary Benefit (INB) and the selected cutpoints. These are controlled by the `what` argument which can be any of `c("nmb", "inb", "cutpoints")`. If a vector is passed, the first value is used. If you choose to visualise the INB, you must your chosen reference strategy for the calculation in the `inb_ref_col`.

```{r, message=FALSE}
plot(sim_screen_obj, what = "nmb")
plot(sim_screen_obj, what = "inb", inb_ref_col = "all")
plot(sim_screen_obj, what = "cutpoints")
```


### Selecting what to show on the plot

The plots show, be default, the median (the dot), the 95% confidence interval (thick vertical lines), and the range (thin vertical lines), and the lines between the points. These can each be shown or hidden independently, and the width of the confidence interval can be controlled using the `ci` argument.

```{r, message=FALSE}
plot(sim_screen_obj)
plot(sim_screen_obj, plot_range = FALSE)
plot(sim_screen_obj, plot_ci = FALSE, plot_range = FALSE)
plot(sim_screen_obj, plot_ci = FALSE, plot_range = FALSE, plot_line = FALSE)
plot(sim_screen_obj)
```


### Other plot modifications

Currently, the lines and dots overlap. We can use `dodge_width` to apply a horizontal dodge for all layers.

```{r, message=FALSE}
plot(sim_screen_obj)
plot(sim_screen_obj, dodge_width = 0.01)
```

The cutpoint methods can be renamed or removed. To rename them, pass a named vector to the `rename_vector` argument. The names of the vector are the new names and the values are the names you wish to replace.

```{r, message=FALSE}
plot(sim_screen_obj)
plot(
  sim_screen_obj, 
  rename_vector = c("Treat All" = "all", "Treat None" = "none")
)
```

You can reorder the methods by passing the order as the `methods_order` argument. Also note that this will remove all methods which aren't included, and it will factor the names AFTER it has renamed them, so if you are both renaming and re-ordering, you must provide the updated names.

```{r, message=FALSE}
plot(sim_screen_obj)
plot(sim_screen_obj, methods_order = c("all", "none"))
plot(
  sim_screen_obj,
  rename_vector = c("Treat All" = "all", "Treat None" = "none"),
  methods_order = c("Treat All", "Treat None")
)
```

The transparency of all layers can be controlled with `plot_alpha`.

```{r, message=FALSE}
plot(sim_screen_obj)
plot(sim_screen_obj, plot_alpha = 0.2)
plot(sim_screen_obj, plot_alpha = 1)
```

## Plotting the results from `do_nmb_sim()`

Many of the same arguments that we used above can be used with the object returned from `do_nmb_sim()`

```{r, include=FALSE}
do_nmb_sim_obj <- sim_screen_obj$simulations[[1]]
```

```{r, eval=FALSE}
do_nmb_sim_obj <- do_nmb_sim(
  n_sims = 500,
  n_valid = 10000,
  sim_auc = 0.7,
  event_rate = 0.1,
  fx_nmb_training = get_nmb_sampler_training,
  fx_nmb_evaluation = get_nmb_sampler_evaluation,
  cutpoint_methods = c("all", "none", "youden", "value_optimising")
)
```

The plots here show the results of a single simulation and compare the available cutpoints.

```{r}
plot(do_nmb_sim_obj)
```

The y-axis variable and names and orders of methods can be controlled in the same way as previously:

```{r}
plot(do_nmb_sim_obj, what = "nmb")

plot(
  do_nmb_sim_obj, 
  what = "inb", 
  inb_ref_col = "all", 
  rename_vector = c("Value-Optimising" = "value_optimising", 
                    "Treat-None" = "none",
                    "Youden Index" = "youden")
)

plot(
  do_nmb_sim_obj, 
  what = "cutpoints", 
  methods_order = c("all", "none", "value_optimising")
)
```

These plots display the median as the solid bar, the grey part of the distributions are the outer 5% of the simulated values and the light blue region is the 95% CI. The fill colours of the histogram are controlled with `fill_cols` and the line for the median is controlled with `median_line_col`. The thickness of the median line is controlled with `median_line_size` and it's transparency with `median_line_alpha`.

```{r}
plot(
  do_nmb_sim_obj, 
  fill_cols = c("red", "blue"), 
  median_line_col = "yellow",
  median_line_alpha = 1,
  median_line_size = 0.9
)
```

The `n_bins` argument controls the number of bins used for the histograms and the `label_wrap_width` is the number of characters above which to start a new line for the facet labels. This can be handy when using detailed names for the methods when the font of the label is relatively large compared to the plot.

```{r, fig.height=3, fig.width=6}
plot(
  do_nmb_sim_obj,
  n_bins = 15,
  rename_vector = c("Value- Optimising" = "value_optimising", 
                    "Treat- None" = "none",
                    "Treat- All" = "all",
                    "Youden Index" = "youden"),
  label_wrap_width = 12
)
```

By default there is an additional theme applied to the facets to bring them closer to each other. Replace the `extra_theme` with `NULL` to revert to the default behaviour, or use your own theme here.

```{r}
plot(do_nmb_sim_obj, extra_theme = NULL)

plot(
  do_nmb_sim_obj, 
  extra_theme = theme(
    panel.spacing = unit(0, "lines"),
    axis.text.x = element_blank(),
    strip.background = element_rect(fill = "red")
  )
)
```

